---
title: "DrugSynergy_DND41"
author: "Hui Mao"
date: "5/18/2020"
output: html_document
---


### Getting Started

This R Markdown is used for analyzing drug combination dose–response data. The example dataset is "DND41_Isobologram.csv". There are 5 methods including: 

*  Chou-Talalay method: CI.
*  Loewe: the synergy score quantifies the excess over the expected response if the two drugs are the same compound.
*  HSA: the synergy score quantifies the excess over the highest single drug response.
*  Bliss: the expected response is a multiplicative effect as if the two drugs act independently.
*  ZIP: Zero interaction potency model, the expected response corresponds to an additive effect as if the two drugs do not affect the potency of each other.

Chou-Talalay method uses the code written by Julie, add dose–response curves for the single drugs, fitted by four-parameter logistic models and one isobologram.

Loewe, HSA, Bliss and ZIP use the SynergyFinder R package to generate synergy scores and visualization for dose–response matrix data.


```{r library}
library(MASS)
library(drc)
library(readxl)
library(ggplot2)
library(readxl)
library(synergyfinder)
```

### Chou-Talalay method


```{r chou, echo=FALSE}
x <- read.csv("/Users/Hui/Drug_synergy/DND41_Isobologram.csv", header = TRUE)
x <- x[,1:7]
x[,4:7] <- x[, 4:7] * 100

combination.vib <- x[x[, 2] != 0 & x[,3] != 0, 7]

A1.vib <- c(as.numeric(unlist(x[x[,2] == 0, 4])), 
            as.numeric(unlist(x[x[,2] == 0, 5])), 
            as.numeric(unlist(x[x[,2] == 0, 6])))/100
A1.vib[A1.vib >1] <- 1
A1.conc <- rep(as.numeric(as.character(unlist(x[x[,2] == 0, 3]))), 3)

A2.vib <- c(as.numeric(unlist(x[x[,3] == 0, 4])), 
             as.numeric(unlist(x[x[,3] == 0, 5])), 
             as.numeric(unlist(x[x[,3] == 0, 6])))/100
 
A2.vib[A2.vib > 1] <- 1
A2.conc <- rep(as.numeric(as.character(unlist(x[x[,3] == 0, 2]))), 3)
#Plot
#A1: GSK4716
G<- drm(A1.vib ~ A1.conc, fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
ED(G, c(30, 50, 70), interval = "delta")
modelFit(G)
summary(G)
plot(G, broken=TRUE, bp=1, ylab="Response",
    xlab="GSK", bty="l", type = "all", xlim = c(0, 20000 ), ylim = c(-0.1, 1))

#A2:Dexamethasone
D<- drm(A2.vib ~ A2.conc, fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
ED(D, c(30, 50, 70), interval = "delta")
modelFit(D)
summary(D)
plot(D, broken=TRUE, bp=1, ylab="Response",
     xlab="Dex", bty="l", type = "all", xlim = c(0, 1500 ), ylim = c(-0.1, 1))

#### 3 parameters with lower bound 0 for ED50
Fors.LL.3 <- drm(A1.vib ~ A1.conc, fct = LL.3())
ED(Fors.LL.3, c(30, 50, 70), interval = "delta")
##### DND4 — Kayleigh GSK4716
Dex.LL.3 <- drm(A2.vib ~ A2.conc, fct = LL.3())
ED(Dex.LL.3, c(20, 30, 50, 70, 80), interval = "delta")

##### ##### DND4 — Kayleigh Dex
Fors.dose <- ED(Fors.LL.3, (100 - combination.vib), interval = "delta")
Dex.dose <- ED(Dex.LL.3, (100 - combination.vib), interval = "delta")


isob.data <- cbind( Dex.dose = Dex.dose[,1], Fors.dose = Fors.dose[,1], comb.Dex.dose = x[x[,2] != 0 & x[,3] != 0, 2], comb.Fors.dose = x[x[,2] != 0 & x[,3] != 0, 3])

isob.data <- transform(isob.data, CI= isob.data[,4]/isob.data[,2] + isob.data[,3] /isob.data[,1])

isob.data <- transform(isob.data, normalized.comb.Fors.dose =  comb.Fors.dose / Fors.dose, normalized.comb.Dex.dose = comb.Dex.dose / Dex.dose)

isob.data <- transform(isob.data, IC = gsub("e:1:", "",rownames(isob.data)))
colnames(isob.data)
colnames(isob.data) <- gsub("Fors", "GSK4716", colnames(isob.data))
rownames(isob.data) <- gsub("e:1:", "",rownames(isob.data))

#plot
gg<-ggplot(isob.data, aes(x=normalized.comb.GSK4716.dose, y= normalized.comb.Dex.dose))
gg1 <- gg + geom_point() + labs(title="Isobologram") + coord_fixed(ratio = 1)+ coord_cartesian(xlim=c(0,1.5), ylim=c(0, 1.5)) + geom_abline(intercept = 1, slope = -1, color="blue")
gg1
```


### SynergyFinder
Efficient implementations for all the popular synergy scoring models for drug combinations, including HSA, Loewe, Bliss and ZIP and visualization of the synergy scores as either a two-dimensional or a three-dimensional interaction surface over the dose matrix. 

* Input data
```{r sfmat, echo=FALSE}
SFDND41 <- read_excel("SF_R_DND41.xlsx")
head(SFDND41)   
dim(SFDND41) 
source("Reshape.R")
dose.response.mat <- mReshapeData(SFDND41,
                                 data.type = "viability",
                                 impute = TRUE,
                                 noise = TRUE,
                                 correction = "part")

str(dose.response.mat)
PlotDoseResponse(dose.response.mat)
```







* Drug synergy scoring


Given a reference model specified by the ‘Method’ parameter, an overall synergy score is calculated as the deviation of phenotypic responses compared to the expected values, averaged over the full dose–response matrix. 

The synergy score for a drug combination is averaged over all the dose combination measurments.
synergy maps highlight synergistic and antagonistic dose regions in red and green colors, respectively.

+ Less than -10: the interaction between two drugs is likely to be antagonistic;
+ From -10 to 10: the interaction between two drugs is likely to be additive;
+ Larger than 10: the interaction between two drugs is likely to be synergistic.

* method: Loewe

```{r sfscoreloewe}

synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'Loewe')
PlotSynergy(synergy.score, type = "all")
```
* method: HSA

```{r sfscoreHSA}

synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'HSA')
PlotSynergy(synergy.score, type = "all")
```


* method: Bliss

```{r sfscoreBliss}

synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'Bliss')
PlotSynergy(synergy.score, type = "all")
```

* method: zip

```{r sfscorezip}

synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'ZIP')
PlotSynergy(synergy.score, type = "all")
```

```{r citation, echo=FALSE}
citation("synergyfinder")
citation('drc')
```

