---
title: "DrugSynergy_PA03C"
author: "Hui Mao"
date: "5/19/2020"
output: html_document
---


```{r library}
library(MASS)
library(drc)
library(readxl)
library(ggplot2)
library(readxl)
library(synergyfinder)
```


### Input data
* Dataset: "Isobologram_PA03C.csv", stacked format "SF_R_PA03C.xlsx". 

Combine "Isobologram - PA03C + Hesperadin - Experiment 1.csv" and "Isobologram - PA03C + Hesperadin - Experiment 2.csv",

* Six replicates per condition.

* AZD2014 (2 levels): 0 (N) and 200 (Y).

* Hesperadin (6 levels):  0, 20, 50, 100, 500 and 2000.

```{r data, echo=FALSE}
SFPA03C <- read_excel("SF_R_PA03C.xlsx")
SFPA03C$response[SFPA03C$response>100] <- 100
head(SFPA03C)   
dim(SFPA03C) 

x <- read.csv("Isobologram_PA03C.csv", header = TRUE)
x <- x[,1:7]
x[,4:7] <- x[, 4:7] * 100

combination.vib <- x[x[, 2] != 0 & x[,3] != 0, 7]

#A1: Hesperadin 
A1.vib <- c(as.numeric(unlist(x[x[,2] == 0, 4])), 
            as.numeric(unlist(x[x[,2] == 0, 5])), 
            as.numeric(unlist(x[x[,2] == 0, 6])))/100
A1.vib[A1.vib >1] <- 1
A1.conc <- rep(as.numeric(as.character(unlist(x[x[,2] == 0, 3]))), 3)
#A2: AZD2014
A2.vib <- c(as.numeric(unlist(x[x[,3] == 0, 4])), 
             as.numeric(unlist(x[x[,3] == 0, 5])), 
             as.numeric(unlist(x[x[,3] == 0, 6])))/100
 
A2.vib[A2.vib > 1] <- 1
A2.conc <- rep(as.numeric(as.character(unlist(x[x[,3] == 0, 2]))), 3)
```

### Exploratory data analysis

* Explore with single drug independently.

There exists numerous sigmoid curves of which the log-logistic is the most common one. It is symmetric and the ED50  is a parameter of the model. Several other models exist, particularly, the non symmetric ones where ED50 is not a “natural” parameter. For example the Weibull-1 model.
```{r sfmat, echo=FALSE}

#A1: Hesperadin single drug
H.m0<- drm(A1.vib ~ A1.conc, fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
ED(H.m0, c(30, 50, 70), interval = "delta")
modelFit(H.m0)
summary(H.m0)
plot(H.m0, bp=1, ylab="Response",
    xlab="Dose of Hesperadin", bty="l", type = "all", xlim = c(0, 5000 ), ylim = c(-0.1, 1))

## Comparing log-logistic and Weibull models
H.m1 <- drm(A1.vib ~ A1.conc,  fct = W1.4())
H.m2 <- drm(A1.vib ~ A1.conc, fct = W2.4())
modelFit(H.m0)
modelFit(H.m1)
modelFit(H.m2)
#W1.4 best, but ED50 similar in LL.4 and W1.4()
summary(H.m1)
ED(H.m1, c(30, 50, 70), interval = "delta")

par(mfrow=c(1,1), mar=c(3.2,3.2,.5,.5), mgp=c(2,.7,0))
plot(H.m0, broken=TRUE, xlab="Dose (mM)", ylab="Response", lwd=2, 
     cex=1.2, cex.axis=1.2, cex.lab=1.2, bty="l")
plot(H.m1, add=TRUE, broken=TRUE, lty=2, lwd=2)
plot(H.m2, add=TRUE, broken=TRUE, lty=3, lwd=2)

arrows(3, 7.5, 1.4, 7.5, 0.15, lwd=2)
text(3,7.5, "Weibull-2", pos=4, cex=1.2)

arrows(2.5, 0.9, 5.7, 0.9, 0.15, lwd=2)
text(3,0.9, "Weibull-1", pos=2, cex=1.2)


#A2: AZD single drug
plot(A2.conc, A2.vib)

```
```{r 2 drug model}
mselect(HA.m1, list(LL.5(), LL.4(), LL.3(), LL.2()), linreg=TRUE, icfct=AIC)

```


```{r twodrug}
# two drug dose response
# HA.m1 the first with treatment as an effect 
HA.m1 <-drm(response~conc_c, AZD2014, data = SFPA03C,fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
modelFit(HA.m1)
summary(HA.m1)
ED(HA.m1, c(30, 50, 70), interval = "delta")
plot(HA.m1, broken=TRUE, bp=.1, bty="l",
     ylab="Response",
     xlab="Dose of Hesperadin")

# HA.m2 the second that looks at the whole complete model without treatment AZD2014
HA.m2 <-drm(response~conc_c, data = SFPA03C,fct = LL.4(names = c("Slope", "Lower Limit", "Upper Limit", "ED50")))
modelFit(HA.m2)
summary(HA.m2)

#plot together
par(mfrow=c(1,1), mar=c(3.2,3.2,.5,.5), mgp=c(2,.7,0))
plot(HA.m1, broken=TRUE, bty="l", xlab="Dose of Hesperadin")
plot(HA.m2, col="red", broken=TRUE, add=TRUE, legend=FALSE)
legend(740,89, legend="ALL",col="red",lty=1, box.lty=0, bg="transparent")

#A test for  significance of treatment (if it is a significant effect).
anova(HA.m1, HA.m2) 

#use the fit and calculate the selectivity factor at ED50

EDcomp(HA.m1, c(50, 50),interval="delta")


# use absoluty value
ED(HA.m1, c(20,30,40,50), type="absolute", interval="delta")
EDcomp(HA.m1, c(20,20), type="absolute",interval="delta")
EDcomp(HA.m1, c(30,30), type="absolute",interval="delta")
EDcomp(HA.m1, c(40,40), type="absolute",interval="delta")
EDcomp(HA.m1, c(50,50), type="absolute",interval="delta")
par(mfrow = c(1, 1), mar=c(3.2,3.2,.5,.5), mgp=c(2,.7,0))
plot(HA.m1, broken=TRUE, bp=.1, bty="l",
      ylab="Response",
     xlab="Dose of Hesperadin")
arrows(  0.1, 50, 56.4603, 50, code=0, lty=1, col="red")
arrows(56.4603 , 50, 56.4603,0, code=0, lty=1, col="red")
arrows(14.3058 , 50, 14.3058, 0, code=0, lty=2, col="red")


```


#2*2 factor
```{r anaova}
res.aov3 <- aov(response ~ conc_c + AZD2014, data = SFPA03C)
summary(res.aov3)
TukeyHSD(res.aov3, which = "AZD2014")
plot(res.aov3, 2)

library("ggpubr")
ggboxplot(SFPA03C, x = "conc_c", y = "response", color = "AZD2014",
          palette = c("#00AFBB", "#E7B800"))


```







* Drug synergy scoring


Given a reference model specified by the ‘Method’ parameter, an overall synergy score is calculated as the deviation of phenotypic responses compared to the expected values, averaged over the full dose–response matrix. 

The synergy score for a drug combination is averaged over all the dose combination measurments.
synergy maps highlight synergistic and antagonistic dose regions in red and green colors, respectively.

+ Less than -10: the interaction between two drugs is likely to be antagonistic;
+ From -10 to 10: the interaction between two drugs is likely to be additive;
+ Larger than 10: the interaction between two drugs is likely to be synergistic.



Efficient implementations for all the popular synergy scoring models for drug combinations, including HSA, Loewe, Bliss and ZIP and visualization of the synergy scores as either a two-dimensional or a three-dimensional interaction surface over the dose matrix. 

Warning!
One of the input drugs has three or less doses measured. Therefore, only Bliss and HSA synergy scores can be calculated, since Loewe and ZIP models require multiple doses for fitting dose-response curves of each drug in the combination. Additionaly, the correction parameter was disabled, which also requires dose-response fitting.

* method: Loewe

```{r sfscoreloewe}
source("Reshape.R")
dose.response.mat <- mReshapeData(SFPA03C,
                                 data.type = "viability",
                                 impute = TRUE,
                                 noise = TRUE,
                                 correction = "part")

str(dose.response.mat)
PlotDoseResponse(dose.response.mat)
synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'Loewe')
PlotSynergy(synergy.score, type = "all")
```
* method: HSA

```{r sfscoreHSA}

synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'HSA')
PlotSynergy(synergy.score, type = "all")
```


* method: Bliss

```{r sfscoreBliss}

synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = 'Bliss')
PlotSynergy(synergy.score, type = "all")
```


```{r citation, echo=FALSE}
citation("synergyfinder")
citation('drc')
```

